---
title: "Grey’s Anatomy Graph Analysis"
author: "Lidan Hifi and Hagai Levi"
date: "9 April 2016"
output:
  html_document:
    keep_md: yes
    toc: true
---

This is the third assignment in introduction to data science course at Ben Gurion University, May 2016.

The main idea of this assignment is analyzing networks- basic metrics of graphs and clustering.   In this assignment we'll make some research on Grey's anatomy characters.

```{r global_options, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE) # remove `##` from prints
library(igraph)
setwd("~/Projects/bgu/greys_anatomy")
```

# Loading data
The graph represent relationships between characters in Grey's anatomy.
It is represented by a two-columns csv file, contains [from, to] pairs.
```{r}
ga.data <- read.csv('ga_edgelist.csv', header = TRUE)
g <- graph.data.frame(ga.data, directed = FALSE)

# remove self loops
g <- simplify(g)
```

# Data exploration
We'll start with a basic data exploration in order to understand the given dataset and the structure of the graph.
Basic exploration:
```{r}
head(ga.data)
```

characters (graph vertices):
```{r}
V(g)$name
```

connections between characters:
```{r}
summary(g)
```

graph visualization:
```{r}
# set seed to make the layout reproducible
set.seed(100)
default_layout = layout.fruchterman.reingold(g)
plot(g, layout = default_layout)
```

## Degree score
```{r}
degree.score <- degree(g)
V(g)$color <- "#d08bcd"
plot(g, layout = default_layout, vertex.size=degree.score * 4, asp=FALSE)
degree.score
```

## Closeness centrality
```{r}
g.closeness <- closeness(g)
V(g)$color <- "#d08bcd"
plot(g, layout = default_layout, vertex.size=g.closeness * 4000, asp=FALSE)
g.closeness
which.max(g.closeness)
```

**Torres** is the most central by closeness centrality metric.

## Betweeness centrality
```{r}
g.betweenness <- betweenness(g)
V(g)$color <- "#d08bcd"
plot(g, layout = default_layout, vertex.size=g.betweenness / 5, asp=FALSE)
g.betweenness
which.max(g.betweenness)
```

**Sloan** is the most central by betweenness centrality metric.

## Eigenvector centrality
```{r}
g.eigen_values <- centr_eigen(g)
V(g)$color <- "#d08bcd"
plot(g, layout = default_layout, vertex.size=g.eigen_values$vector * 10, asp=FALSE)
g.eigen_values$vector
V(g)[which.max(g.eigen_values$vector)]
```

**Karev** is the most central by eigenvector centrality metric.

# Community Clustering

## Walktrap clustering
Find clusters using the `walktrap` algorithm.
```{r}
community <- walktrap.community(g)
plot(g, layout=default_layout, vertex.size=5, vertex.color=community$membership, asp=FALSE)

# number of clusters and their size
sizes(community)

# modularity
modularity(community)
```


## Edge-betweenness clustering
Find clusters using the `edge betweenness` algorithm.
```{r}
community <- edge.betweenness.community(g)
plot(g, layout=default_layout, vertex.size=5, vertex.color=community$membership, asp=FALSE)

# number of clusters and their size
sizes(community)

# modularity
modularity(community)
```
